---
title: "Hydraulically-vulnerable trees are buffered by deep-water access in a tropical forest community"
author: ''
date: ''
output:
  bookdown::html_document2:
    number_sections: no
    theme: flatly
    fig_caption: yes
  bookdown::word_document2:
        reference_docx: manuscriptstyle.docx
        fig_caption: yes
  bookdown::pdf_document2: default
bibliography: references.bib
csl: new-phytologist.csl
---

```{r, include=FALSE, echo = FALSE}
if(!require("pacman", quietly=TRUE)) install.packages("pacman")
suppressPackageStartupMessages(
  pacman::p_load(bookdown, knitr, pander, citr, broom))
options(tinytex.verbose = TRUE)
source("code/load.R")
source("code/manuscript_data_figs.R")
```

```{r setup, echo = FALSE}
library(knitr)
opts_knit$set(eval.after = "fig.cap")
```

## Intro

Historic climatic variation is relatively small (Mora et al., 2013), and, though data have been scant, trees' hydraulic safety margins are reportedly relatively narrow (Choat et al., 2012; but see Oliveira et al., 2019). 


## M&M

To estimate effective rooting depth, we tested alternate empirical models for species 5-yrly census growth (Eq. \@ref(eq:M-chosen) and those in SI), which use effective rooting depth as an input. Many intrinsic and extrinsic factors limit plant growth, including soil moisture, VPD, leaf area seasonality and radiation [@Lawrence.2019]. Rhizosphere soil water potential ($\mathrm{\Psi}$~soil~) down-regulates $\mathrm{\Psi}$~leaf~ [@Christoffersen:2016kd2d] such that in moist forests $\mathrm{\Psi}$~leaf~ equilibrates with $\mathrm{\Psi}$~soil~ at night, declining through the day. Using species-specific leaf hydraulic vulnerability curves, $\mathrm{\Psi}$~leaf~ can be converted to $K_\textrm{leaf}$, which in turn controls growth [@Sack.2012]. We therefore used maximum diurnal ($K_\textrm{leaf}$) determined from $\mathrm{\Psi}$~soil~ as a proxy for the soil water limitation on growth. $K_\textrm{leaf}$ in our models thus incorporates the effects of rooting depth and rhizosphere water potential. This approach, however, does not account for the buffering effect of stem water storage capacitance on $\mathrm{\Psi}$~leaf~, which may cause the model to slightly overestimate ERD (see Discussion). VPD affects growth non-linearly, with growth increasing with VPD upto a VPD threshold, then declining as leaf pores (stomata) close [@Grossiord.2020]. Apart from stomatal control, leaf-drop may further limit water-uptake and growth. We therefore regressed species-specific growth against multiplicative or additive effects of one or more factors calculated daily and averaged over 5-yrly censuses: VPD, $K_\textrm{leaf}$ and the leaf area index (LAI), all standardized between 0 and 1. Based on our preliminary analyses, additional incorporation of radiation did not improve model fitting. At our study site, due to the coincidence of higher temperature, higher radiation, and lower humidity during the dry season, the impact of VPD may as well be a combined impact of not only temperature and humidity, but also radiation.


For each species $s$ we calculated mean mortality rate across all intervals as $\textrm{Mort}_{s,\textrm{mean}} = \frac{1}{6}\sum_{t=1}^{6} {\textrm{Mort}_{s,t}}$. 

Briefly, this isotopic approach makes use of a naturally formed gradient in deuterium to hydrogen ratio of soil water vertically belowground, which is generally formed due to evaporative enrichment (here, during the dry season). Shallow soil water has a greater deuterium to hydrogen content than deep soil water (here, $\mathrm{\delta}$^2^H~soil~ ranged from -16.5 to -50 ‰ over the depth of 0.5-1 m), with ground-water having the lowest deuterium to hydrogen content (here, -60 ‰), typically reflecting un-fractionated rain water. As root water-uptake is generally a non-fractionating process, $\mathrm{\delta}$^2^H~xylem~ of a tree reflects the signature of the source water. Thus water-sourcing depth of the tree can be identified as the depth with the same value of $\mathrm{\delta}$^2^H~soil~, or as that of $\mathrm{\delta}$^2^H~xylem~, with the uncertainty that the value of $\mathrm{\delta}$^2^H~xylem~ may also result from a mixing of water from shallower or deeper depths .

## Results

The hydraulic efficiency-safety trade-off between high values of maximum stem conductivity (high $K_\textrm{max,stem}$) and cavitation-resistance (low $\Psi$~50,stem~ or $\Psi$~88,stem~) is hypothesized although community-wide examples are rare [@Hacke:2006; @Gleason.2016]. We report such a trade-off ($n$ = 26 species) at our study site stem level. 

Species with deeper ERD showed lower mean mortality rates across the 35-yrs record (1982-2015; *R*^2^ = 0.11, *p* = 0.075, *n* = `r length(erd.mrate.sp)`; Fig. \@ref(fig:erd-mort)), and this relationship was stronger among evergreen species. 

Specifically for the water environment (or, resource), observations of individual trade-offs between environment-trait, trait-trait and trait-demography has identified an acquisitive-conservative resource economic spectrum similar to those for light and nutrients, although, for water, examples of all trade-offs observed simultaneously are rare, and below-ground traits or environment linkages are weak [@Reich:2014; @Sterck:2011]. Acquisitive species have traits with high capacity to acquire, move, store and use water (higher hydraulic conductivity, low wood density) and achieve high growth, but build flimsy tissues (high specific leaf area) and are less tolerant of low resource availablity, whereas species with conservative traits are better protected from carbon losses and are more physiologically drought tolerant (greater embolism resistance, larger safety margins). Whether and how the acquisitive-conservative spectrum holds true for the water dynamics vertically below-ground is not clear. Acquisitive species are spatially associated with wet sites (high resource) such as valleys or those with high rainfall [@Kursar:2009; @Oliveira:2019]. Deep rooting may be expected to allow an acquisitive strategy by enabling a steady water supply. However, the degree to which deep rooting is associated with an acquisitive strategy may depend on the site-specific dynamics of water availability by depth, which is driven by precipitation regime and hydrology. For example, in pulse-driven systems *shallow*-rooted plants are associated with acquisitive traits (higher hydraulic conductivity) and opportunistic strategies (e.g. dry deciduousness) that maximise the use of the water-pulse, while *deep*-rooted plants show conservative traits (cavitation-resistant xylem, osmoregulation) [@Schwinning:2013 and references therein]. 

\
At the community level, we filled data gaps in $LMA$ for `r sum(c.lma.gaps[-which(names(c.lma.gaps) == "original")])` species (of total `r sum(c.lma.gaps)`)  using linear relationships with other variables: `r c.lma.gaps[[which(names(c.lma.gaps) == "filled-LMA-DISC")]]` species with $LMA_{disc}$, `r c.lma.gaps[[which(names(c.lma.gaps) == "filled-LMA-Nobby")]]` with $LMA_{lamina}$, and `r c.lma.gaps[[which(names(c.lma.gaps) == "filled-LEAF")]]` with $LMA_{leaf}$, defined as the mean leaf mass per unit area measured for the entire leaf including the petiole (g m^-2^) for leaves receiving direct sunlight ($\overline{LMA} = `r leaf.cf[1]` + `r leaf.cf[2]`LMA_{leaf}$; Adj. *R*^2^ = `r leaf.r2`; *P* `r leaf.p`, *n* = `r leaf.n`). For the `r sum(c.lma.gaps)` species thus with $LMA$, we gap-filled data on  WSG for `r sum(c.wsg.gaps[-which(names(c.wsg.gaps) == "original")])` species using a linear relationship with wood specific gravity reported by @Chave:2009 ($\overline{WSG} = `r chave.cf[1]` + `r chave.cf[2]`WSG_{chave}$; Adj. *R*^2^ = `r chave.r2`; *P* `r chave.p`, *n* = `r chave.n`).


#Discussion

Mortality rates of shallow ERD species may have been exacerbated by increased drought events during the study period. @Condit.2017 showed that half of the extreme dry seasons (< 1 SD annual water deficit) during 1931-2015 (85-yr) occurred during 1982-2005 (24-yr) of the 34-yr long study period (1982-2015).


Our result of deep rooted species being more hydraulically vulnerable is consistent with observations for a seasonal forest in the Amazon . @Johnson:2018fad found that mortality decreased with deep rooting even though deep rooted species were more vulnerable to xylem cavitation (high $\mathrm{\Psi}$~88,stem~) and had narrower safety-margins ($\mathrm{\Psi}$~min~ - $\mathrm{\Psi}$~88,stem~). @Nardini:2015 found that a species sourcing deep water in a Karst system showed lower hydraulic failure and mortality in an extreme drought despite having high $\mathrm{\Psi}$~88,stem~ and narrower safety margin compared to two species sourcing shallow-water.

at Mudumalai is nearly one (Penman-Monteith PET = 980 $\pm$ 64 mm)
and precipitation to potential evapotranspiration ratio at Mudumalai is nearly one (Penman-Monteith PET = 980 $\pm$ 64 mm), whereas at BCI it is nearly two (Penman-Monteith PET = `r round(mean(clim$pet.PM))` $\pm$ `r round(sd(clim$pet.PM))` mm;`r min(clim$Year)`-`r max(clim$Year)`; this study, @Guo.2016),

\
\   We found a strong scaling relationship between commonly collected 'soft' traits at the community scale (LMA and WSG) and parameters of leaf vulnerability (or percent loss of conductance, PLC) curves, a 'hard' trait, based on one of the largest dataset for the tropics ($n$ = `r length(sp.exp.param$sp)`). This allowed us to scale-up the leaf vulnerability curves at the community scale. We find that leaf conductance declined more rapidly with declining $\mathrm{\Psi}$~leaf~ for species with higher values of WSG and LMA (Fig. S\@ref(fig:obs-plc-curves)). Whether such scaling relationships between hard and soft traits hold true for other forests, and their wider applicability, remains to be seen as data on PLC curves are limited in tropical forests, but supportive and counter examples both exist [@McCulloh.2011; @Hoffmann.2011]. WSG alone explained `r round(b.wsg.r2*100,0)`% of the variance in $B$ ($p$ `r  b.wsg.p`). 

However, trees' stem diameter change is a sum total of not just growth, the addition of wood, but also changes due to reversible stem dehydration (xylem, phloem and bark), the importance of which increases with temporal resolution [@Chitra-Tarak.2016; @Chitra-Tarak.2015]. 

Use of dendrometer band data would thus entail accounting for reversible stem dehydration

This would also open the possibility of testing, whether deciduous species are more shallow rooted than evergreen species as generally observed [@Fan:2017]

, and suggests that greater emphasis should be given to predictive modeling of tree root allocation and distribution (e.g. @Wijk.2001) as well as trees' hydrological environments


\
\ Finally, rooting depths may be parametarized within DGVMs such as ELM-FATES using plant hydrodynamic schemes [@Christoffersen:2016kd2d; @Koven.2020], simultaneously modeling plant water-availability with plant water-extraction. In general, including differences in tree rooting depths in ELM-FATES would improve understanding of plant competition for water (and other resources) and more accurately represent plant's drought strategies, leading to better predictions of existing and future forest function and composition.

The model in @Chitra-Tarak.2018 estimated six species-specific phenomenological parameters: three as root fraction for corresponding lumped vadose zone layers, and three as shape parameters for water uptake vs availability curves. For each species, they calculated daily water uptake successively from three layers, that is, the supply, to satisfy the demand of the daily maximum evapotranspiration, a function of PET and LAI. They regressed total supply to total demand ratio over census intervals as a growth coefficient against growth (five data points), and chose values for six parameters that best-fit (highest *R*^2^, that is, two additional parameters). Our inverse ERD model fits only two parameters by way of regressing depth-specific model against five growth data points. Our model was constrained on VPD and species leaf vulnerability curves, variables that are shown to be physiologically meaningful.

The model in CT2018 were not corroborated with such data as existing methods were inadequate to extract water bound to the saprolite found in their dry forest study site (Jean Riotte, pers. comm.). The model of below-ground water-availability in this study (ELM-FATES) differs from the conceptual water balance model used by CT2018 and was calibrated with data on long-term soil moisture dynamics at a point location (0-0.15m) and plot-wide measurements by depth (0.15, 0.4 and 1 m), as well as, evapotranspiration and stream discharge. The model in CT2018 was calibrated only on stream discharge and water table levels.
